Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-02-03 parts/sa.tex
add new extraction algorithm

This section gives a recursive method to find a scenario that concretizes
a given objective sequence $\w \in \OS$.
All the definitions above are well-defined provided that
$\uconcr(\w)\neq\emptyset$,
which can be demonstrated by \pref{th:approxinf} or \pref{thm:ordered-ua}.
The justification of these definitions can be found by analysing the demonstrations
of these theorems.

\newcommand{\res}{\mathbf{res}}
\newcommand{\osres}{\mathbf{OSres}}
\newcommand{\autoBS}{\mathbf{autoBS}}
\newcommand{\rec}{\mathbf{rec}}
% \newcommand{\thisobj}[1]{P_{#1}}
% \newcommand{\thisseq}[1]{\Theta_{#1}}
\newcommand{\thisobj}{P}
\newcommand{\thisseq}{\Theta}
\newcommand{\R}{R}

First, $\osres$ is used to split the objective sequence $\w$ to solve,
by considering successive objectives, in the fashion of \pref{thm:ordered-ua}.
This function calls $\res$ which allows to solve an objective $P$
by either re-target it (if $\gCont(\PHsort(P), P) \neq \emptyset$)
of by searching for a solution to this objective in $\BS(P)$.
This search may require to try several objective sequences $\zeta \in \BS(P)$.
For each trial, the chosen objective sequence $\zeta$ is solved with $\rec$,
which recursively created a scenario to play the whole sequence.
This scenario is itself recursively created by combining
smaller scenarios allowing to reach each required hitter of each action
with the function $\thisseq$.
Finally, $\thisseq$ calls $\res$ anew to reach the hitters
that are necessarily contained in other automata.
In the end, a scenario can be obtained by analysing all values of
$\osres_s(\w)$ (\pref{def:concret})
for all possible initial state $s \subseteq \ctx$ in the considered context,
and all possible bounce sequences $\zeta \in \BS(P)$ that are chosen during the resolution.

\begin{definition}[$\osres : \PHl \times \OS \to \Sce$]
\label{def:concret}
For all $s \in \PHl$ and $\w \in \OS$, we define:
\[
  \osres_s(\w) \DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\w = \emptyseq$} \\
      \res_s(\w_1) \concat
      \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
        & \text{ otherwise} \\
    \end{cases}
\]
and, for all $s \in \PHl$ and $P \in \Obj$:
\[\res_s(P) \DEF
  \begin{cases}
    \varepsilon
      & \text{if } \target{P} = \bounce{P} \\
    \osres_s(P^Q \concat Q)
      & \text{if } \exists Q \in \Obj, (P, Q) \in \Bee{\Obj}{\Obj} \\
    \rec_s(\zeta)
      & \text{otherwise, with } \zeta \in \BS(P)
  \end{cases}
\]
with $P^Q = \PHobj{\PHtarget(P)}{\PHtarget(Q)}$.
Furthermore:
\begin{align*}
  \forall \zeta \in \BS,
  \rec_s(\zeta) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\zeta = \emptyseq$} \\
      \thisseq_s(\zeta_1) \concat \zeta_1 \concat
      \rec_{s \play \thisseq_s(\zeta_1) \concat \zeta_1}
      (\zeta_2 \concat \dots \concat \zeta_{\card{\zeta}})
        & \text{ otherwise}
    \end{cases} \\
  \forall h \in \PHh,
  \thisseq_s(h) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\hitter{h} = \emptyset$} \\
      \R_s(h_1) \concat \thisseq_{s \play \R_s(h_1)}(h_2 \concat \dots \concat h_{\card{h}})
        & \text{ otherwise}
    \end{cases} \\
  \forall a_i \in \Proc,
  \R_s(a_i) &\DEF \res_s(\PHobj{\PHget{s}{\PHsort(a_i)}}{a_i})
\end{align*}
and for all action $h \in \PHh$, $(h_i)_{i \in \segm{1}{\card{\hitter{h}}}}$
is a sequence containing all hitters of $h$ in an arbitrary order.
\end{definition}

>>>>

This section gives a recursive method to find a scenario that concretizes
the reachability of a given local state $\myp \in \Proc$,
from a given initial state $s_0 \in \PHl$,
provided that \pref{th:approxinf} answered positively
for a context $\ctx \in \Ctx$ so that $s_0 \subset \ctx$.
The algorithm proposed in the following relies on a traversal of the Local Causality Graph
$\cwB$ that has been used for this conclusion.
The correctness of this extraction can be demonstrated
in the same fashion as \pref{th:approxinf}.

This algorithm consists in visiting all required nodes in a given order,
to build a sequence of actions that concretizes the objective.
The actions to perform on a given node,
listed below,
depend on the current state, the type of node, the markers on this node
and the state of the traversal which can be either “descending” (D) or “ascending” (A).
Nodes in $\Sol$ can be marked with a sequence of actions
and nodes in $\sSol$ can be marked with a set of local states.
The traversal starts in node $\myp$ in descending mode,
the starting state is $s = s_0$
and the initial output is the empty sequence $\emptyseq$.

\begin{itemize}
  \item In a node $a_k \in \Proc$:
    \begin{itemize}
      \item[D)] Descend in the node $\PHobj{\PHget{s}{a}}{a_k} \in \Obj$.
      \item[A)] Ascend in the parent $\sSol$ node, if any,
        and remove the element $a_k$ from its marking set.
        If the node has no $\sSol$ parent, the traversal is complete
        and the output of the algorithm is the current output.
    \end{itemize}

  \item In a node $\PHobj{a_j}{a_k} \in \Obj$:
    \begin{itemize}
      \item[D)] Descend in an arbitrarily chosen $\sSol$ node,
        and mark it with a sequence $\zeta \in \BS$ so that $\abstr{\zeta} = pps$.
      \item[A)] Ascend in the parent $\Proc$ or $\Obj$ node.
    \end{itemize}

  \item In a node $pps \in \Sol$ with the current marking $\zeta$:
    \begin{itemize}
      \item If $\zeta \neq \emptyseq$,
        descend in the node $ps \in \sSol$
        so that $ps = \hitter{\zeta_1}$
        and mark it with the set $ps$.
      \item If $\zeta = \emptyseq$,
        ascend in the parent $\Obj$ node and carry on ascending.
    \end{itemize}

  \item In a $\sSol$ node with the current marking $m$:
    \begin{itemize}
      \item If $m \neq \emptyset$,
        descend in a child node $a_k \in \Proc$ arbitrarily chosen,
        so that $a_k \in m$ and carry on descending.
      \item If $m = \emptyset$,
        ascend in the parent $\sSol$ node.
        Let $\zeta$ be the current marking of this node.
        If $\zeta_1$ is playable in $s$:
          append $\zeta_1$ to the output,
          change the current state to $s \play \zeta_1$
          and mark this node with $\zeta_{2..|\zeta|}$.
        If $\zeta_1$ is not playable
        (which means that $\target{\zeta_1}$ has changed):
          go to the node $\PHobj{\PHget{s}{\PHsort(\bounce{\zeta_1})}}{\bounce{\zeta_1}} \in \Obj$
          and start descending.
    \end{itemize}
\end{itemize}

% * On applique les règles suivants de parcours où, à tout moment, s représente l'état courant, et l'état du parcours peut être « en descente » ou « en montée » :
% 
% - Dans un nœud a_k \in LS :
%   - Si en cours de descente, descendre dans le nœud s[a] />* a_k \in Obj.
%   - Si en cours de remontée, monter dans le nœud Sol parent et enlever l'élément a_k de son marquage.
% 
% - Dans un nœud a_j />* a_k \in Obj :
%   - Si en cours de descente, descendre dans un nœud pps \in NSol au choix. Marquer ce nœud avec une séquence \zeta \in BS tq \zeta^ = pps.
%   - Si en cours de remontée, monter dans le nœud LS ou Obj parent.
% 
% - Dans un nœud pps \in NSol (marqué par \zeta) :
%   - Si \zeta ≠ \emptyseq, descendre dans le nœud ps \in Sol tq ps = hitters(\zeta_1). Marquer le nœud ps par l'ensemble ps.
%   - Si \zeta = \emptyseq, alors remonter dans le nœud Obj parent et continuer la remontée.
% 
% - Dans un nœud Sol marqué par m :
%   - Si m ≠ \emptyset, alors descendre dans un des nœuds fils a_k \in LS tq a_k \in m.
%   - Si m = \emptyset, alors remonter dans le nœud NSol parent, marqué par \zeta. Si \zeta_1 est jouable, jouer cette action et marquer le nœud par \zeta_{2..|\zeta|}. Si \zeta_1 n'est pas jouable (c.-à-d. si target(\zeta_1) a changé), aller au nœud s[sort(bounce(\zeta_1))] />* bounce(\zeta_1) \in Obj et commencer une descente.

%%%%%%%%% Tnetative 1 %%%%%%%%%%%%%%%%%%%%%
% 
% \newcommand{\res}{\mathbf{res}}
% \newcommand{\osres}{\mathbf{OSres}}
% \newcommand{\autoBS}{\mathbf{autoBS}}
% \newcommand{\rec}{\mathbf{rec}}
% % \newcommand{\thisobj}[1]{P_{#1}}
% % \newcommand{\thisseq}[1]{\Theta_{#1}}
% \newcommand{\thisobj}{P}
% \newcommand{\thisseq}{\Theta}
% \newcommand{\R}{R}
% 
% First, $\osres$ is used to split the objective sequence $\w$ to solve,
% by considering successive objectives, in the fashion of \pref{thm:ordered-ua}.
% This function calls $\res$ which allows to solve an objective $P$
% by either re-target it (if $\gCont(\PHsort(P), P) \neq \emptyset$)
% of by searching for a solution to this objective in $\BS(P)$.
% This search may require to try several objective sequences $\zeta \in \BS(P)$.
% For each trial, the chosen objective sequence $\zeta$ is solved with $\rec$,
% which recursively created a scenario to play the whole sequence.
% This scenario is itself recursively created by combining
% smaller scenarios allowing to reach each required hitter of each action
% with the function $\thisseq$.
% Finally, $\thisseq$ calls $\res$ anew to reach the hitters
% that are necessarily contained in other automata.
% In the end, a scenario can be obtained by analysing all values of
% $\osres_s(\w)$ (\pref{def:concret})
% for all possible initial state $s \subseteq \ctx$ in the considered context,
% and all possible bounce sequences $\zeta \in \BS(P)$ that are chosen during the resolution.
% 
% \begin{definition}[$\osres : \PHl \times \OS \to \Sce$]
% \label{def:concret}
% For all $s \in \PHl$ and $\w \in \OS$, we define:
% \[
%   \osres_s(\w) \DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\w = \emptyseq$} \\
%       \res_s(\w_1) \concat
%       \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
%         & \text{ otherwise} \\
%     \end{cases}
% \]
% and, for all $s \in \PHl$ and $P \in \Obj$:
% \[\res_s(P) \DEF
%   \begin{cases}
%     \varepsilon
%       & \text{if } \target{P} = \bounce{P} \\
%     \osres_s(P^Q \concat Q)
%       & \text{if } \exists Q \in \Obj, (P, Q) \in \Bee{\Obj}{\Obj} \\
%     \rec_s(\zeta)
%       & \text{otherwise, with } \zeta \in \BS(P)
%   \end{cases}
% \]
% with $P^Q = \PHobj{\PHtarget(P)}{\PHtarget(Q)}$.
% Furthermore:
% \begin{align*}
%   \forall \zeta \in \BS,
%   \rec_s(\zeta) &\DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\zeta = \emptyseq$} \\
%       \thisseq_s(\zeta_1) \concat \zeta_1 \concat
%       \rec_{s \play \thisseq_s(\zeta_1) \concat \zeta_1}
%       (\zeta_2 \concat \dots \concat \zeta_{\card{\zeta}})
%         & \text{ otherwise}
%     \end{cases} \\
%   \forall h \in \PHh,
%   \thisseq_s(h) &\DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\hitter{h} = \emptyset$} \\
%       \R_s(h_1) \concat \thisseq_{s \play \R_s(h_1)}(h_2 \concat \dots \concat h_{\card{h}})
%         & \text{ otherwise}
%     \end{cases} \\
%   \forall a_i \in \Proc,
%   \R_s(a_i) &\DEF \res_s(\PHobj{\PHget{s}{\PHsort(a_i)}}{a_i})
% \end{align*}
% and for all action $h \in \PHh$, $(h_i)_{i \in \segm{1}{\card{\hitter{h}}}}$
% is a sequence containing all hitters of $h$ in an arbitrary order.
% \end{definition}

%%%%%%%% Tentative 2 %%%%%%%%%%%%%%%%%%%%%%%
% % \newcommand{\resP}{\mathbf{\Obj res}}
% % \newcommand{\resOS}{\mathbf{\OS res}}
% \newcommand{\resP}{\mathsf{res}^\Obj}
% \newcommand{\resOS}{\mathsf{res}^\OS}
% 
% \newcommand{\res}{\mathbf{res}}
% \newcommand{\autoBS}{\mathbf{autoBS}}
% \newcommand{\rec}{\mathbf{rec}}
% 
% Pour tout objectif $P$ dans $\cwB$,
% on note $X_P$ une séquence d'objectifs obtenue en effectuant
% un parcours en profondeur d'abord de $\cwB$.
% Un tel parcours existe car $\cwB$ ne comporte pas de cycle.
% 
% $$\res_s(\w) = \resOS_s((X_{\w_i})_{i \in \indexes{\w}})$$
% 
% \[
%   \resOS_s(\w) \DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\w = \emptyseq$} \\
%       \resP_s(\w_1) \concat
%       \resOS_{s \play \resP_s(\w_1)}(\w_{2..\card{\w}})
%         & \text{ otherwise} \\
%     \end{cases}
% \]
% 
% $
% \resP_s(\PHobj{a_j}{a_k}) \DEF
%   \todo{resP}$
% 
% % $\resP_s(P) \DEF
% %     \begin{cases}
% %       \emptyseq
% %         & \text{ if $P$ is trivial} \\
% %       \res_s(\w_1) \concat
% %       \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
% %         & \text{ otherwise} \\
% %     \end{cases}$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
